Optimizing Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD Transient

### 2019-10-16 paper/cvpr2020/main.tex
pass over simulated results section
added paper to discuss
David's comments.

@@

%%%%%%%%% TITLE
\title{Optimizing Monocular Depth Estimation with Global Depth Histogram Matching using a Single SPAD}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

>>>>

% TODO:
%@article{caramazza2018neural,
%  title={Neural network identification of people hidden from view with a single-pixel, single-photon detector},
%  author={Caramazza, Piergiorgio and Boccolini, Alessandro and Buschek, Daniel and Hullin, Matthias and Higham, Catherine F and Henderson, Robert and Murray-Smith, Roderick and Faccio, Daniele},
%  journal={Scientific reports},
%  volume={8},
%  number={1},
%  pages={11945},
%  year={2018},
%  publisher={Nature Publishing Group}
%}
%

@@

%%%%%%%%% TITLE
\title{Optimizing Monocular Depth Estimation with Global Depth Histogram
  Matching using a Single SPAD Transient}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org
\end{document}
}
}

### 2019-10-16 paper/cvpr2020/sections/method.tex
David's comments.

Consider a diffused laser that emits a pulse at time $t = 0$ with time-varying intensity $g(t)$ illuminating some 3D scene. We parameterize the geometry of the scene as a depth map $z(x, y)$, where each of the 3D points has also some unknown reflectivity $\alpha$ at the wavelength of the laser. Ignoring interreflections of the emitted light within the scene, a single-pixel diffused SPAD then integrates all the light scattered back from the scene towards the detector as
%
\begin{equation}
	s \left( t \right)= \int_{\Omega_x} \int_{\Omega_y} \frac{\alpha \left( x,y \right)}{z(x,y)^2}  g \left( t - \frac{2z(x,y)}{c} \right) dx dy ,
	\label{eq:pulse_integral} 
\end{equation}  
%
where $c$ is the speed of light and $\Omega_{x,y}$ is the angular range of the diffuser. In this formulation, we assume that the diffuser spreads light uniformly over all angles. Each time such a light pulse is emitted into the scene and scattered back to the detector, the single-pixel SPAD time-stamps up to one of the returning photons with some probability. This is not necessarily the first returning photon. The process is repeated millions of times per second with the specific number of emitted pulsed being controlled by the repetition rate of the laser, which is typically adjusted to match the deadtime of the SPAD~\cite{Heide:2018}. Each detected photon arrival event is discretized into a histogram $h$ of the form
%
\begin{equation}
  h[n] \sim \mathcal{P} \left( \eta \int_{n\Delta t}^{(n+1)\Delta t} \left(f * s \right) \left( t \right)  dt + b \right),	
	\label{eq:spad_measurements}
\end{equation}
%
where $(n\Delta t, (n+1) \Delta t)$ models the $n^{th}$ time interval of the
temporal histogram, $\eta$ is the photon detection probability of the SPAD, $f$
is a function that models the temporal uncertainty in the detector, and $b$
represents ambient background light and falsely detected photons known as dark
count. As derived in previous work, the combination of signal, background, and
noise can be modeled as an inhomogeneous Poisson process
$\mathcal{P}$~\cite{Kirmani:2014,Shin2015}. Finally, as in \cite{Xin2019} we adopt the term
\textit{transient} for our histogram $h[n]$.

\subsection{Ambient Rejection and Falloff Correction}
To remove the ambient and dark count photons from the histogram, we apply a
preprocessing step. Assuming that the first $N$ bins of the SPAD histogram
contain only ambient and dark count photons (this is equivalent to assuming the
minimum image depth is larger than $Nc\Delta T/2$), we estimate 
and dark count rate $b$ as $\hat b = \frac{1}{N}\sum_{n=0}^N h[n]$ We then
note that that shining a laser through a 2D diffuser and
measuring a single pixel is very physically similar to recording a single 
non-line-of-sight measurement. We
compute the absolute first difference $d[n] = \abs{h[n] - h[n+1]}$ and find 
the set of edges. Intuitively, given a sufficiently large value of $b$, the Poisson sample
$h[n] \sim \mathcal{P}(b)$ should behave roughly like a Gaussian with variance $b$, and the difference
$h[n] - h[n+1]$ should behave like a Gaussian with variance $2b$. Therefore, we
can detect edges that occur between bins that contain actual signal photons and
bins that do not by computing the set 
\begin{equation}
  \mathcal{E} = \mset{n}{d[n] > \beta\sqrt{2\hat b}}.
  \label{eq:edge_set}
\end{equation}
In this work, we use $\beta = 5$.
We identify the closest and furthest edges $n_{first} = \min \mathcal{E}$ and
$n_{last} = \max \mathcal{E}$.
From \cite{Xin2019} Proposition 3, these edges
correspond to the discontinuties in the depth map 
at the closest and furthest image depths, and therefore all of the signal
photons must be in the bins between these two edges.
We refine these estimates by
computing $n'_{first} = \max\set{n_{first}}\cup\mset{n}{h[n] > \hat b, n \leq n_{first}}$ and
$n'_{last} = \min\set{n_{last}}\cup\mset{n}{h[n] > \hat b, n \geq n_{last}}$
Finally, we reject all photons outside of the valid range by clamping those bins
to 0, and we subtract our ambient estimate from the photons within the valid
range, giving an ambient-free transient
\begin{equation}
  h'_{clean}[n] = \begin{cases}
    0 & n < n'_{first} \\
    \max(h[n] - \hat b, 0) & n'_{first} \leq n \leq n'_{last} \\
    0 & n > n'_{last} \\
  \end{cases}.
  \label{eq:h_clean}
\end{equation}
To compensate for the radiometric falloff, we scale each bin by its squared depth,
which is consistent with our hardware prototype's radiometric falloff model.
This gives
\begin{equation}
  h_{clean}[n] = h'_{clean}[n] \cdot r_n^2
  \label{eq:h_scaled}
\end{equation}
where $r_n= \paren*{n + \frac12}\paren*{\frac{c\Delta t}{2}}$ is the depth
corresponding to bin $n$
As a final preprocessing step, we re-bin the transient, applying the Spacing-Increasing Discretization
(SID) of \cite{Fu2018} to produce $h_{target}$. This reduces the number of bins in the histogram
matching procedure, giving speedups while maintaining performance. In this work,
we use 140 bins for the simulated results and 600 bins for the captured results.

@@

\subsection{Exact Histogram Matching}
% \begin{algorithm}[H]
%  \caption{Exact Histogram Matching} 
%  \label{alg:ehm}
%  \input{sections/figures/exact_hist_matching_pseudo.tex}
% \end{algorithm}
Exact histogram matching is a procedure that takes a source image and a target
histogram and maps the pixel values of the source image so that the image's
histogram matches the target histogram. Given a single RGB image, we pass it
through an MDE to produce a depth map $\hat z_0 \in [\ell, u]^{M \times N}$
where $\ell$ and $u$ are the lower and upper bounds of the depth range of
interest, respectively. In a parallel processing step,
we compute an estimate of the scene albedo using an intrinsic imaging method and
take the color channel of the resulting albedo image that corresponds to the
wavelength of the laser illumination. This gives a reflectance estimate $W \in
[0, 1]^{M \times N}$. We
then compute a weighted histogram $h_{source}$of the depth map $\hat z_0$ where the
per-pixel weights are given by $W$, and the bin edges are given by the SID
discretization algorithm of \cite{Fu2018}.

We match the histogram $h_{source}$ to $h_{target}$ using the method of
\cite{Morovic2002}. We compute a pixel movement
matrix $T$ such that $T_{mn}$ is the amount of mass to move from $h_{source}[m]$
to $h_{target}$. We do so using the procedure outlined in \ref{alg:pixel_movement}

\begin{algorithm}[H]
 \caption{Pixel Movement} 
 \label{alg:ehm}
 \begin{algorithmic}
  \Procedure{FindMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array of zeros.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{\ell=1}^{n-1} T[m, \ell]$
        \State $p_t \gets \sum_{\ell=1}^{m-1} T[\ell, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}

Finally, once we have obtained $T$, for a pixel $z_0[i, j]$ in depth bin $k$, we sample the new depth bin from the
distribution $T[k, :]/\sum_{n=1}^NT[k,n]$ (which is a distribution over target
depth bins $1,\ldots,N$) to get the bin of $\hat z[i,j]$. We can then apply the
standard SID decoding procedure to get the depth corresponding to that bin
\textcolor{red}{Does the above seem reasonable to include? Not sure if the
  pseudocode is helpful/necessary or not.}
%\subsection{Implementation Details}
%For the Monocular Depth Estimator, we use pretrained versions of the
%the Deep Ordinal Regression Network (DORN) \cite{} and the DenseDepth Network.
%The exact histogram matching method is as described in \cite{}.

>>>>

Consider a diffused laser that emits a pulse at time $t = 0$ with time-varying
intensity $g(t)$ illuminating some 3D scene. We parameterize the geometry of the
scene as a distance map $z(x, y)$, where each of the 3D points has also some
unknown reflectivity $\alpha$ at the wavelength of the laser. Ignoring
interreflections of the emitted light within the scene, a single-pixel diffused
SPAD integrates the light scattered back from the scene towards the
detector as
%
\begin{equation}
	s \left( t \right)= \int_{\Omega_x} \int_{\Omega_y} \frac{\alpha \left( x,y \right)}{z(x,y)^2}  g \left( t - \frac{2z(x,y)}{c} \right) dx dy ,
	\label{eq:pulse_integral} 
\end{equation}  
%
where $c$ is the speed of light and $\Omega_{x,y}$ is the angular range of the
diffuser. In this formulation, we assume that the diffuser spreads light
uniformly over all angles. Each time such a light pulse is emitted into the
scene and scattered back to the detector,
the single-pixel SPAD time-stamps up to one of the returning photons with some
probability. The process is
repeated millions of times per second with the specific number of emitted pulses
being controlled by the repetition rate of the laser.
Each detected
photon arrival event is discretized into a histogram $h$ of the form
%
\begin{equation}
  h[n] \sim \mathcal{P} \left( \eta \int_{n\Delta t}^{(n+1)\Delta t} \left(f * s \right) \left( t \right)  dt + b \right),	
	\label{eq:spad_measurements}
\end{equation}
%
where $(n\Delta t, (n+1) \Delta t)$ models the $n^{th}$ time interval of the
temporal histogram, $\eta$ is the photon detection probability of the SPAD, $f$
is a function that models the temporal uncertainty in the detector, and $b$
represents ambient background light and false positive detections known as \textit{dark
count}. As derived in previous work, the combination of signal, background, and
noise can be modeled as an inhomogeneous Poisson process
$\mathcal{P}$~\cite{Kirmani:2014,Shin2015}. Finally, as in \cite{Xin2019}, we adopt the term
\textit{transient} for the histogram $h[n]$.

\subsection{Ambient Rejection and Falloff Correction}
Before histogram matching can be applied, the raw SPAD transient is processed in
three stages.

In the first stage, we remove the ambient and dark count photons from the transient. 
We compute an estimate of $b$ by assuming that the first $N$ bins of the SPAD transient
contain only ambient and dark count photons (this is equivalent to assuming the
minimum image distance is larger than $Nc\Delta T/2$) as:
\begin{equation}
  \hat b = \frac{1}{N}\sum_{n=0}^N h[n]. 
  \label{eq:ambient_estimate}
\end{equation}
We identify the first and last bins $n_{first}$ and $n_{last}$
corresponding to the minimum and maximum scene depths. We do this with a simple
edge thresholding procedure, computing the absolute first difference $d[n] =
\abs{h[n] - h[n+1]}$. If $h[n]$ and $h[n+1]$ are sampled according to $\mathcal{P}(b)$ and $b$ is large, then 
$h[n] - h[n+1]$ is well-approximated with a Gaussian with mean $0$ and variance $2b$.
Therefore, we can extract a candidate edge set $\mathcal{E}$ as:
\begin{equation}
  \mathcal{E} = \mset{n}{d[n] > \beta\sqrt{2\hat b}}.
  \label{eq:edge_set}
\end{equation}
In this work, we use $\beta = 5$, as it gives good results across both simulated
and captured data.
We compute initial estimates $n'_{first} = \min \mathcal{E}$ and
$n'_{last} = \paren*{\max \mathcal{E}} + 1$ respectively. By appealing to Proposition 3 in
\cite{Xin2019} and noting the similarity of our approach to collecting a single
non-line-of-sight SPAD measurement we can conclude that our estimates must be
close to the true values of $n_{first}$ and $n_{last}$.
We refine these estimates by walking them outwards until their bin values drop
below the ambient estimate $\hat b$, resulting in
$\hat n_{first} = \max\paren*{\set{n'_{first}}\cup\mset{n}{h[n] > \hat b, n \leq n'_{first}}}$ and
$\hat n_{last} = \min\paren*{\set{n'_{last}}\cup\mset{n}{h[n] > \hat b, n \geq n'_{last}}}$.
Finally, we reject all photons outside of the valid range by clamping those bins
to 0, and we subtract our ambient estimate from the photons within the valid
range:
\begin{equation}
  h'_{clean}[n] = \begin{cases}
    0 & n < \hat n_{first} \\
    \max(h[n] - \hat b, 0) & \hat n_{first} \leq  n \leq \hat n_{last} \\
    0 & n > \hat n_{last} \\
  \end{cases}.
  \label{eq:h_clean}
\end{equation}

In the second stage, we set 
\begin{equation}
  h_{clean}[n] = h'_{clean}[n] \cdot r_n^2
  \label{eq:h_scaled}
\end{equation}
where $r_n= \paren*{n + \frac12}\paren*{\frac{c\Delta t}{2}}$ is the depth
corresponding to bin $n$. This compensates for the radiometric falloff in a way
that is consistent with our hardware prototype's falloff model

In the final stage, we re-bin the transient, applying the Spacing-Increasing Discretization
(SID) of \cite{Fu2018} to produce $h_{target}$. This discretization scheme is
exponential rather than linear, allocating more bins to closer distances and fewer
bins to farther distances. The bin edges $t_i$ are given by the following equation,
parameterized by the number of bins $K$ and the range of distances $[\ell, u]$: 
\begin{equation}
  t_i = e^{\log(\ell) + \frac{\log(u/\ell) \cdot i}{K}}  \qquad i = 0,\ldots, K.
  \label{eq:sid_bin_edges}
\end{equation}
This reduces the number of bins in the histogram
matching procedure, giving speedups while maintaining performance. For the
simulated results,
we use $K = 140$ bins, and $(\ell, u) = (0.657, 9.972)$, (these numbers are
inspired by the discretization scheme for the original DORN network in
\cite{Fu2018})
 For the indoor caputred results, we use $K = 600$ bins, and $(\ell, u)=(0.4, 9) + \epsilon$ 
and for the outdoor captured results, we use $K=600$ and $(\ell, u)=(0.4, 11) +
\epsilon$ where $\epsilon \approx 0.094$ is the $z$ offset between the
SPAD and the Kinect (see Hardware for details).

@@

\subsection{Histogram Matching}
% \begin{algorithm}[H]
%  \caption{Exact Histogram Matching} 
%  \label{alg:ehm}
%  \input{sections/figures/exact_hist_matching_pseudo.tex}
% \end{algorithm}
Histogram matching is a procedure that takes a source image and a target
histogram and maps the pixel values of the source image so that the image's
histogram matches the target histogram. Given a single RGB image, we pass it
through an MDE to produce a depth map. In a parallel processing step,
we compute an estimate $W$ of the scene albedo using an intrinsic imaging method and
take the color channel of the resulting albedo image that corresponds to the
wavelength of the laser illumination.
We then compute a weighted histogram $h_{source}$ of the depth map $\hat z_0$ where the
per-pixel weights are given by $W$, and the bin edges are given by the SID
discretization algorithm in \cite{Fu2018}.

We match the histogram $h_{source}$ to $h_{target}$ using the method of
\cite{Morovic2002}. We compute a pixel movement
matrix $T$ such that $T[m, n]$ is the amount of mass that must be moved from $h_{source}[m]$
to $h_{target}[n]$. We do so using the procedure outlined in \ref{alg:ehm}

\begin{algorithm}[H]
 \caption{Pixel Movement} 
 \label{alg:ehm}
 \begin{algorithmic}
  \Procedure{FindMovement}{$h_s$ of length $M$, $h_t$ of length $N$} 
    \State Initialize $T$ as an $M \times N$ array of zeros.
    \For{$m$ in $1,\ldots,M$}
      \For{$n$ in $1,\ldots,N$}
        \State $p_s \gets \sum_{i=1}^{n-1} T[m, i]$
        \State $p_t \gets \sum_{i=1}^{m-1} T[i, n]$
        \State $T[m, n] \gets \min(h_s[m] - p_s, h_t[n] - p_t)$
      \EndFor
    \EndFor
    \State \textbf{return} $T$
  \EndProcedure
 \end{algorithmic}
\end{algorithm}

Finally, once we have obtained $T$, for a pixel $z_0[i, j]$ in bin $k$, we sample the new bin from the
distribution $T[k, :]/\sum_{n=1}^NT[k,n]$ (which is a distribution over target
depth bins $1,\ldots,N$) to get the bin of $\hat z[i,j]$. We can then apply the
standard SID decoding procedure to get the depth corresponding to that bin
\textcolor{red}{Does the above seem reasonable to include? Not sure if the
  pseudocode is helpful/necessary or not.}
%\subsection{Implementation Details}
%For the Monocular Depth Estimator, we use pretrained versions of the
%the Deep Ordinal Regression Network (DORN) \cite{} and the DenseDepth Network.
%The exact histogram matching method is as described in \cite{}.
